\ttitle

### 2013-10-10 Chapters/Chapter2/temporal_segmentation.tex
Remove note
working on outline of proposed method

\begin{itemize}
  \item Given overview of segmentation techniques, for times series data
  \item Use different ``point-of-views'', or terms
  \item ``Segmentation''
  \item ``Change detection''
  \item ``Novelty detection''
  \item Specific view on \gls{svm}s
\end{itemize}

This section gives an overview of the literature on temporal segmentation in the context of human activity recognition.
It takes a look on different implementations and methodologies.
A wide range of terms and subtle differences are used in the field, such as `segmentation', `change detection', `novelty detection' and `outlier detection'.
These will be the categorical terms for which we discuss the literature.
Finally we will discuss other applications of \gls{svm}s in the context of these terms.

@@

The time series data from a sensor can be considered as a stochastic process *** is this a correct term? ***.
Probabilistic models can be constructed on that signal, yielding in probabilistic and Bayesian based segmentation methods.
The \gls{cusum}-methods takes a statistical approach and relies on the log-likelihood ratio \cite{gustafsson1996marginalized} to measure the difference between two distributions.
To calculate the ratio, the probability density functions need to be calculated.
The method of Kawahara \etal \cite{kawahara2009change} proposes to estimate the ratio of probability densities (known as the \emph{importance}), based on the log likelihood of test samples, directly, without explicit estimation of the densities.
The method by Liu \etal \cite{liu2013change} uses a comparable dissimilarity measure using the \gls{kliep} algorithm.
They claim this results in a robuster approach for real-world scenarios.
Although this is a model-based method, no strong assumptions (parameter settings) are made on the models.

>>>>

\begin{itemize}
  \item Given overview of segmentation techniques, for times series data
  \item Use different ``point-of-views'', or terms
  \item ``Segmentation''
  \item ``Change detection''
  \item ``Novelty detection''
  \item Specific view on \glspl{svm}
\end{itemize}

This section gives an overview of the literature on temporal segmentation in the context of human activity recognition.
It takes a look on different implementations and methodologies.
A wide range of terms and subtle differences are used in the field, such as `segmentation', `change detection', `novelty detection' and `outlier detection'.
These will be the categorical terms for which we discuss the literature.
Finally we will discuss other applications of \glspl{svm} in the context of these terms.

@@

The time series data from a sensor can be considered as being drawn from a certain stochastic process.
Probabilistic models can be constructed on that signal, yielding in probabilistic and Bayesian based segmentation methods.
The \gls{cusum}-methods takes a statistical approach and relies on the log-likelihood ratio \cite{gustafsson1996marginalized} to measure the difference between two distributions.
To calculate the ratio, the probability density functions need to be calculated.
The method of Kawahara \etal \cite{kawahara2009change} proposes to estimate the ratio of probability densities (known as the \emph{importance}), based on the log likelihood of test samples, directly, without explicit estimation of the densities.
The method by Liu \etal \cite{liu2013change} uses a comparable dissimilarity measure using the \gls{kliep} algorithm.
They claim this results in a robuster approach for real-world scenarios.
Although this is a model-based method, no strong assumptions (parameter settings) are made on the models.

### 2013-10-10 Chapters/Chapter3/one_class_svm.tex
working on outline of proposed method

This minimization problem can be solved (using \gls{qp}) and transformed to its Lagrange dual formulation.
To do so, the Lagrange multipliers $a_i >= 0$ are introduced and the decision function becomes:
\begin{equation}\label{eq:svm_lagrange}
  f(x) = \operatorname{sgn}( \sum_{i=1}^n \alpha_i y_i K(x, x_i) + b),
\end{equation}
where $K(x, x_i) = \phi(x)^T\phi(x_i)$ (which is further discussed in section \ref{subsec:kernel_trick}).
Here every data point $\mathcal{I}$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
Since it is shown that under certain circumstances \gls{svm}s show an equality to sparse representations \cite{girosi1998equivalence,smola1998connection}, there will often be relatively few Lagrange multipliers with a non-zero value.

@@

We will discuss two implementations of \gls{oc-svm}s.
The first is the \gls{nu-svm} by Sch\"olkopf \etal \cite{scholkopf1999support} and closely follows the above problem statement regarding density estimation.
The second, on which we will base our change detection method, is the \gls{svdd} by Tax and Duin \cite{tax2004support}.

@@

\subsection{\acrlong{nu-svm}s}\label{subsec:nu-svm}
The first of the \gls{oc-svm} methods we will discuss is often referred to as \gls{nu-svm} and introduced by Sch\"olkopf \etal \cite{scholkopf1999support}.
Instead of estimating the density function of an distribution $P$, it focuses on an easier problem: the algorithm find regions in the input where the ``probability density lives''.
This results in a function such that most of the data is in the region where the function is nonzero.

The constructed decision function $\mathcal{F}$ resembles the function discussed in Section \ref{subsec:svm}.
It returns the value $+1$ in a (possibly small) region capturing most of the data points, and $-1$ elsewhere.
The method maps the data points from input space $\mathcal{I}$ to a feature space $\mathcal{F}$ (following classical \gls{svm}s).
In that space $\mathcal{F}$ it separates the data points with maximal margin from the origin, with a hyperplane.
For a new data points $x$, the function value $f(x)$ determines wheter the data point is part of the distribution (\ie the value is $+1$) or a novelty (\ie the value is $-1$).
The hyperplane is represented by $g(x) = w \cdot \phi(x) + \rho = 0$ and the decision function is $f(x) = \operatorname{sgn}(g(x))$.
This hyperplane and the separation from the origin is illustrated in Figure \ref{fig:nu-svm}.

@@

In the classical \gls{svm} objective function, as denoted in Equation \ref{eq:svm_objective}, the parameter $C$ decided the smoothness of the boundary, with respect to the slack variables $\xi_i$.
In the formulation of \gls{nu-svm} the equivalent parameter is $\nu \in (0,1)$ (hence the name).
It characterizes the solution in two ways:
\begin{enumerate}
  \item $\nu$ is an upper bound on the fraction of outliers, \ie training examples regarded as out-of-class.
  \item $\nu$ is a lower bound on the fraction of \gls{sv}s, \ie training examples with a nonzero Lagrange multiplier $\alpha_i$.
\end{enumerate}
When $\nu$ approaches $0$, the penalty factor for nonzero Lagrange multipliers ($\frac{1}{\nu n}$) becomes infinite, and thus the solution resembles a \emph{hard margin} boundary.

>>>>

This minimization problem can be solved (using \gls{qp}) and transformed to its Lagrange dual formulation.
To do so, the Lagrange multipliers $a_i >= 0$ are introduced and the decision function becomes:
\begin{equation}\label{eq:svm_lagrange}
  f(x) = \operatorname{sgn}( \sum_{i=1}^n \alpha_i y_i K(x, x_i) + b),
\end{equation}
where $K(x, x_i) = \phi(x)^T\phi(x_i)$ (which is further discussed in section \ref{subsec:kernel_trick}).
Here every data point $\mathcal{I}$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
Since it is shown that under certain circumstances \glspl{svm} show an equality to sparse representations \cite{girosi1998equivalence,smola1998connection}, there will often be relatively few Lagrange multipliers with a non-zero value.

@@

We will discuss two implementations of \glspl{oc-svm}.
The first is the \gls{nu-svm} by Sch\"olkopf \etal \cite{scholkopf1999support} and closely follows the above problem statement regarding density estimation.
The second, on which we will base our change detection method, is the \gls{svdd} by Tax and Duin \cite{tax2004support}.

@@

\subsection{\acrlong{nu-svm}}\label{subsec:nu-svm}
The first of the \gls{oc-svm} methods we will discuss is often referred to as \gls{nu-svm} and introduced by Sch\"olkopf \etal \cite{scholkopf1999support}.
Instead of estimating the density function of an distribution $P$, it focuses on an easier problem: the algorithm find regions in the input where the ``probability density lives''.
This results in a function such that most of the data is in the region where the function is nonzero.

The constructed decision function $\mathcal{F}$ resembles the function discussed in Section \ref{subsec:svm}.
It returns the value $+1$ in a (possibly small) region capturing most of the data points, and $-1$ elsewhere.
The method maps the data points from input space $\mathcal{I}$ to a feature space $\mathcal{F}$ (following classical \glspl{svm}).
In that space $\mathcal{F}$ it separates the data points with maximal margin from the origin, with a hyperplane.
For a new data points $x$, the function value $f(x)$ determines wheter the data point is part of the distribution (\ie the value is $+1$) or a novelty (\ie the value is $-1$).
The hyperplane is represented by $g(x) = w \cdot \phi(x) + \rho = 0$ and the decision function is $f(x) = \operatorname{sgn}(g(x))$.
This hyperplane and the separation from the origin is illustrated in Figure \ref{fig:nu-svm}.

@@

In the classical \gls{svm} objective function, as denoted in Equation \ref{eq:svm_objective}, the parameter $C$ decided the smoothness of the boundary, with respect to the slack variables $\xi_i$.
In the formulation of \gls{nu-svm} the equivalent parameter is $\nu \in (0,1)$ (hence the name).
It characterizes the solution in two ways:
\begin{enumerate}
  \item $\nu$ is an upper bound on the fraction of outliers, \ie training examples regarded as out-of-class.
  \item $\nu$ is a lower bound on the fraction of \glspl{sv}, \ie training examples with a nonzero Lagrange multiplier $\alpha_i$.
\end{enumerate}
When $\nu$ approaches $0$, the penalty factor for nonzero Lagrange multipliers ($\frac{1}{\nu n}$) becomes infinite, and thus the solution resembles a \emph{hard margin} boundary.

### 2013-10-10 Chapters/Chapter3/outlier_detection.tex
working on outline of proposed method

\subsection{M-estimators}
*** Move this section to somewhere in the end; is specification/extension, not base material ***
To make a method less sensitive to outliers (in the training data) techniques from robust statistics are applied.
The term \emph{robustness} has many interpretations, one of them that it ``signifies insensi­tivity to small deviations from the [prior] assumptions [about the underlying situation]'', according to Huber \cite{huber2009robust}.
Methods have been proposed to apply robust statistics in the form of M-estimators to \gls{svm}s, such as \cite{choi2009least,chen2004m,suykens1999least}.

>>>>

\subsection{M-estimators}
*** Move this section to somewhere in the end; is specification/extension, not base material ***
To make a method less sensitive to outliers (in the training data) techniques from robust statistics are applied.
The term \emph{robustness} has many interpretations, one of them that it ``signifies insensi­tivity to small deviations from the [prior] assumptions [about the underlying situation]'', according to Huber \cite{huber2009robust}.
Methods have been proposed to apply robust statistics in the form of M-estimators to \glspl{svm}, such as \cite{choi2009least,chen2004m,suykens1999least}.

### 2013-10-10 Chapters/Chapter4/change_indication.tex
working on outline of proposed method

% !TEX root = ../../main.tex
\section{Change Indication}\label{sec:method_change_indication}

>>>>

% !TEX root = ../../main.tex
\section{Change Indication}\label{sec:method_change_indication}
*** TODO: change section title ***
*** TODO: ***
\emph{Follow methodology of ``A unifying framework for detecting outliers and change points from time series'' \cite{takeuchi2006unifying}.
It creates a two-stage process of first searching for outliers, and then using the ``outlier-score'' to find (sudden) change points, by a weighted average of a moving window.
That eventual score can be thresholded (as in the paper) or processed with something like CUSUM (proposal).
Looks like my proposed methods, in that is combines outliers and gives a score to change points.}

The proposed method of this thesis follows the unifying framework as introduced by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and an similar implementation by Camci \cite{camci2010change} with \glspl{svm}.
The unifying framework combines the detection of outliers with change points and divides it in two stages.
The first stage determines the outliers in a time series by giving a score based on the deviation from a learned model, and thereby creates a new time series.
The second stage runs on that new created time series and calculates a average over a window of the outlier scores.
The problem of change detection is then reduced to outlier detection over that average-scored time series.
This method is named \gls{changeFinder} by the authors.
The implementation by Camci, which uses \glspl{svm} to detect changes is named \acrlong{svcpd}.

The problem statement and formal definition, following Tackeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change} is the following.
The algorithm needs to find \emph{sudden} changes in the time series data.
In other words, slowly changing properties in the data are not considered to be changes.
This is in line with the search of changes in activities, since we are only interested in different activities (which are represented by sudden changes) instead of changes within an activity.
Considered a time series $x_1 x_1 \dots$, which is drawn from a stochastic process $p$.
Each $x_t$ (t = 1, 2, \dots) is a $d$-dimensional real valued vector and $p$ a probability density function of the sequence $x_1 x_2 \dots$.
Assume $p$ can be decomposed in two different \gls{iid} stationary stochastic processes $p^1$ and $p^2$ and are one-dimensional Gaussian density functions.
For a time point $a$ data points for which $t < a$ are drawn from $p^1 = N(\mu_1, \sigma_1^2)$ and for $t \geq a$ from $p^2 = N(\mu_2, \sigma_a^2)$.
If $p^1$ and $p^2$ are different, then the time point $t = a$ is a \emph{change point}.
In \cite{takeuchi2006unifying} the similarity between the stochastic processes are expressed by the \gls{kliep} divergence $D(p^2||p^1)$.
The problem with this measure is that, as the authors conclude and Camci discusses, it is not able to detect a change by decrease in variance.

Whereas \gls{changeFinder} uses double probability estimation algorithm, our approach follows \gls{svcpd} by constructing a \gls{svm} over a sliding window.
The \gls{svcpd} algorithm uses the location of new data points in the feature space $\mathcal{F}$ with respect to the hypersphere and the hypersphere's radius $R$ to determine whether the new data point represents a change point.

*** TODO: What is my contribution compared to Camci? Distance of outliers to hypersphere? ***

### 2013-10-10 glossary-acronyms.tex
wip
Work on mathematical formulas
Add paper to Appendix C
WIP: chapter 4, description of experiment methods
Add stuff on Scholkopf one-svm
Complete part on v-SVM
Literature review, temporal segmentation
working on outline of proposed method

@@

@@

@@

@@

@@

>>>>

\newacronym{aic}{AIC}{Akaike Information Criterion}

\newacronym{bic}{BIC}{Bayesian Information Criterion}

@@

\newacronym{hmm}{HMM}{Hidden Markov Model}

@@

\newacronym{iid}{i.i.d.}{independently and identically distributed}

\newacronym{kliep}{KLIEP}{Kullback-Leibler Importance Estimation Procedure}

\newacronym{mle}{MLE}{Maximum Likelihood Estimates}

@@

\newacronym{nu-svm}{$\nu$-SVM}{$\nu$-Support Vector Machine}

\newacronym{occ}{OCC}{One-Class Classifier}

\newacronym{oc-svm}{OC-SVM}{One-Class Support Vector Machine}

\newacronym{rbf}{RBF}{Radial Base Function}

\newacronym{roc}{ROC}{Receiver-Operating Characteristic}

\newacronym{pca}{PCA}{Principal Component Analysis}

\newacronym{rt}{RT}{Ratio-Thresholding}

@@

\newacronym{qp}{QP}{Quadratic Programming}

\newacronym{sv}{SV}{Support Vector}

\newacronym{svc}{SVC}{Support Vector Classifier}

\newacronym{svcpd}{SVCPD}{Support Vector based Change Detection}

@@

\newacronym{swab}{SWAB}{Sliding Window And Buttom-up}
